### 2017-02-12 real/draft.tex
Section on Poisson-Gamma convolution

We return to the general formulation of \refeq{sll} and re-arrange its
factors into
\begin{align}
  \lleq{asc}
  p(Q\cond n,\bml)
  &= \int \rmdx{\lambda}\; \rmdx{\bmphi}\;
    p(\lambda\cond n)\;
    p(\bmphi\cond n,\bml)\;
    p(Q\cond \lambda,\bmphi), \\
  \lleq{asd}
  p(Q\cond \lambda,\bmphi)
  &= \sum_N
    p(N\cond \lambda)\;
    \int \rmdx{\bmL}\;
    p(Q\cond N,\bmL)\;
    p(\bmL\cond N,\bmphi),
\end{align}
with $p(\bmL\cond N,\bmphi) = \prod_{j=1}^N p(L_j\cond\bmphi)$ as
before. Again all quantities are related to a single bin in frequency.
Writing the moments and variance of $L$ as\footnote{The $k$-th moment
  of variable $x$ is defined as the expectation value $\mu_k =
  \expect{x^k} = \int \rmdx{x}\,x^k\,p(x),\; k = 0,1,2,\ldots$.} %
\begin{align}
  \lleq{ram}
  \mu =  \mu_1 &\equiv \expect{\Lum}, \qquad
  \mu_2 \equiv \expect{\Lum^2}, \\
  \sigma^2 &\equiv \variance{\Lum}
  = \expect{\Lum^2} - \expect{\Lum}^2,
\end{align}
it is easy to show that, for the Poissonian compound process
(\ref{asd}), the first moment and variance of $Q$ are related to those
of $L$ by
%%%%%%%%%%%%%%%%
\fredmargin{citation} %
%%%%%%%%%%%%%%%
\begin{align}
  \lleq{raa}
  \expect{Q} = \lambda \mu, \qquad
  \variance{Q} = \lambda \mu_2 = \lambda (\mu^2 + \sigma^2) \,.
\end{align}
for any probability $p(L\cond\bmphi)$.
%
Higher-order quantities such as the skewness and kurtosis are
suppressed by powers of $\lambda^{-1/2}$.\footnote{ The variance is
  the second so-called \textit{cumulant}. The relation for all orders
  is that the $k$-th order cumulant $\kappa_q(Q)$ is equal to
  $\lambda$ times the $k$-th order moment of $L$, $\mu_k$. The
  skewness of $p(Q\cond \lambda,\bmphi)$ is therefore
  $\kappa_3(Q)/\kappa_2(Q)^{3/2} = \lambda \mu_3/(\lambda \mu_2)^{3/2}
  = \lambda^{-1/2} \mu_3/\mu_2^{3/2}$.}  %
%%%%%%%%%%%
\hans{The magnitude of the correction to Eq.~(\ref{rab}) due to
  $\kappa_3$ and higher orders is unclear.}
%%%%%%%%%%%
From Theorem 4.3.1 of Ref.~\cite{bening2002generalized}, we conclude
that the asymptotic distribution of $\Lumtot$ when
$\lambda \to \infty$, given the first and second moment of $\Lum$, is
a Gaussian or normal distribution,
\begin{align}
  \lleq{rab}
  p(\Lumtot \cond \lambda,\bmphi)
  &= \GaussianDist \left( \Lumtot \cond \expect{Q}, \variance{Q} \right) %
    = \GaussianDist \left( \Lumtot \cond \lambda \mu, \lambda (\mu^2 {+} \sigma^2) \right).
\end{align}
This drastic simplification eliminates both the sum over $N$ and the
integral over $\bmL$ in \refeq{asd}.

@@

@@

@@

>>>>

We return to the general formulation of \refeq{sll} and re-arrange its
factors into
\begin{align}
  \lleq{asc}
  p(Q\cond n,\bml)
  &= \int \rmdx{\lambda}\; \rmdx{\bmphi}\;
    p(\lambda\cond n)\;
    p(\bmphi\cond n,\bml)\;
    p(Q\cond \lambda,\bmphi), \\
  \lleq{asd}
  p(Q\cond \lambda,\bmphi)
  &= \sum_N
    p(N\cond \lambda)\;
    \int \rmdx{\bmL}\;
    p(Q\cond N,\bmL)\;
    p(\bmL\cond N,\bmphi),
\end{align}
with $p(\bmL\cond N,\bmphi) = \prod_{j=1}^N p(L_j\cond\bmphi)$ as
before. Again all quantities are related to a single bin in frequency.
Writing the moments and variance of $L$ as\footnote{The $k$-th moment
  of variable $x$ is defined as the expectation value $\mu_k =
  \expect{x^k} = \int \rmdx{x}\,x^k\,p(x),\; k = 0,1,2,\ldots$.} %
\begin{align}
  \lleq{ram}
  \mu =  \mu_1 &\equiv \expect{\Lum}, \qquad
  \mu_2 \equiv \expect{\Lum^2}, \\
  \sigma^2 &\equiv \variance{\Lum}
  = \expect{\Lum^2} - \expect{\Lum}^2,
\end{align}
it is easy to show that, for the Poissonian compound process
(\ref{asd}), the first moment and variance of $Q$ are related to those
of $L$ by
%%%%%%%%%%%%%%%%
\fredmargin{citation} %
%%%%%%%%%%%%%%%
\begin{align}
  \lleq{raa}
  \expect{Q} = \lambda \mu, \qquad
  \variance{Q} = \lambda \mu_2 = \lambda (\mu^2 + \sigma^2) \,.
\end{align}
for any probability $p(L\cond\bmphi)$.
%
Higher-order quantities such as the skewness and kurtosis\footnote{
  The variance is the second so-called \textit{cumulant}. The relation
  for all orders is that the $k$-th order cumulant $\kappa_q(Q)$ is
  equal to $\lambda$ times the $k$-th order moment of $L$,
  $\mu_k$. The skewness of $p(Q\cond \lambda,\bmphi)$ is therefore
  $\kappa_3(Q)/\kappa_2(Q)^{3/2} = \lambda \mu_3/(\lambda \mu_2)^{3/2}
  = \lambda^{-1/2} \mu_3/\mu_2^{3/2}$.}  %
are suppressed by powers of $\lambda^{-1/2}$.
%%%%%%%%%%%
\hans{H comment (not for text): The magnitude of the correction to
  Eq.~(\ref{rab}) due to $\kappa_3$ and higher orders is unclear.}
%%%%%%%%%%%
From Theorem 4.3.1 of Ref.~\cite{bening2002generalized}, we conclude
that the asymptotic distribution of $\Lumtot$ when
$\lambda \to \infty$, given the first and second moment of $\Lum$, is
a Gaussian or normal distribution,
\begin{align}
  \lleq{rab}
  p(\Lumtot \cond \lambda,\bmphi)
  &= \GaussianDist \left( \Lumtot \cond \expect{Q}, \variance{Q} \right) %
    = \GaussianDist \left( \Lumtot \cond \lambda \mu, \lambda (\mu^2 {+} \sigma^2) \right).
\end{align}
This drastic simplification eliminates both the sum over $N$ and the
integral over $\bmL$ in \refeq{asd}.

@@

\subsection{Poisson-gamma convolution}\label{sec:poisson-gamma}

\hans{H: This subsection is temporary. It is motivated by the question
  whether the use of a gaussian for the single-packet luminosity
  $p(L_j\cond\bmphi)$ in (\ref{rae}) is necessary and whether making
  it a gaussian is a source of bias since $L_j$ cannot be
  negative. The original choice of a Gamma distribution as in
  (\ref{spc}) is more appropriate and could eliminate one source of
  systematic bias. Below, I have derived a closed form for
  $p(Q\cond \lambda,\bmphi)$ using the gamma-distributed
  $p(L_j\cond \alpha,\beta)$ of (\ref{spc}). The final result
  (\ref{ffin}) below could be used to replace the gaussian result
  (\ref{rab}) for insertion into (\ref{asd}). Inclusion of the entire
  derivation below is not envisaged; rather, we would include only
  salient issues and equation in the text itself. For the moment, the
  question is what Fred has to say about this version ie whether it
  can be implemented numerically and how it compares to the above
  results.

  The limit of this section to Gaussian can be investigated later if necessary.
}

We start again from (\ref{asd}) and use (\ref{spc})
\begin{align}
  \lleq{asdd}
  p(Q\cond \lambda,\bmphi)
  &= \sum_N
    p(N\cond \lambda)\;
    \int \rmdx{\bmL}\;
    p(Q\cond N,\bmL)\;
    p(\bmL\cond N,\bmphi)\\
  \lleq{spcd}
  p(L_j\cond \alpha,\beta) 
  &= \frac{\beta^\alpha L_j^{\alpha-1}e^{-\beta L_j}}{\Gamma(\alpha)}.
\end{align}
The moment generating function of $p(L_j\cond \alpha,\beta)$ is
\begin{align}
  M[t\cond p(L_j\cond \alpha,\beta) ]
  &= E[e^{tL_j}] = \int_0^\infty dL_j e^{tL_j}\,p(L_j\cond \alpha,\beta) 
    = (1 - t/\beta)^{-\alpha}
\end{align}
The convolution of $N$ Gamma distributions yields the moment generating
function
\begin{align}
  M[t\cond p(Q\cond \alpha,\beta) ]
  &= (1-t/\beta)^{-N\alpha}
\end{align}
and so we obtain the well-known result that the convolved distribution
is also a Gamma distribution with shape parameter $N\alpha$,
\begin{align}
  p(Q\cond \alpha,\beta) 
  &= \frac{\beta^{N\alpha} Q^{N\alpha-1}e^{-\beta Q}}{\Gamma(N\alpha)}.
\end{align}
Inserting this into the $N$-sum, we have
\begin{equation}
  \lleq{gms}
  \boxed{
  p(Q\cond \lambda,\alpha,\beta) 
  = \frac{e^{-\lambda - \beta Q}}{Q}
    \sum_{N=0}^\infty 
    \frac{[\lambda \beta^\alpha Q^\alpha]^N}{N!\;\Gamma(N\alpha)}.
    }
\end{equation}
This series itself should converge quickly because of the
$N!\;\Gamma(N\alpha)$ in the denominator. It is possible to find an
analytic solution for some special cases of $\alpha$.
\begin{itemize}
\item For $\alpha=1$ we have
  \begin{align}
    p(Q\cond \lambda,\alpha{=}1,\beta) 
    &= \frac{e^{-\lambda - \beta Q}}{Q}
      \sum_{N=0}^\infty 
      \frac{[\lambda \beta Q]^N}{N!\;(N-1)!}.
  \end{align}
  Comparing the above to the series definition of the modified Bessel function,
  \begin{align}
    I_1(z) &= \sum_{k=0}^\infty \frac{(z/2)^{2k+1}}{k!\;(k+1)!},
  \end{align}
  we can identify $z = 2\sqrt{\lambda\beta Q}$ to get
  \begin{equation}
    \boxed{
      p(Q\cond \lambda,\alpha{=}1,\beta) 
      = \sqrt{\frac{\lambda\beta}{Q}}  e^{-\lambda - \beta Q} 
      \;I_1\!\left(2\sqrt{\lambda\beta Q}\right)
      }
  \end{equation}
  as derived by Fisher and Cornish (1960). We often prefer to use the 
  corresponding hypergeometric function formulation, in terms of which
  \begin{align}
    p(Q\cond \lambda,\alpha{=}1,\beta) 
    &= \lambda\beta\; e^{-\lambda - \beta Q} 
      \;{}_0F_1(2; \lambda\beta Q).
  \end{align}

\item For integer $\alpha = 2,3,\ldots$, we rely on a paper by Withers and
  Nadarajah (2011) from which we get an expression for the (p,q) hypergeometric 
  function, defined as
  \begin{align}
    {}_pF_q\left( a_1,a_2,\ldots, a_p\;;\; b_1,b_2,\ldots,b_q\;;\; z \right)
    &= \sum_{k=0}^\infty \frac{z^k}{k!} \;
      \frac{(a_1)_k(a_2)_k\cdots (a_p)_k}{(b_1)_k(b_2)_k\cdots (b_q)_k}
  \end{align}
  in terms of the Pochhammer symbol $(a)_k = \Gamma(a+k)/\Gamma(a)$. Withers
  and Nadarajah found that (and I checked)
  \begin{align}
    (N\alpha)! 
    &= (\alpha^\alpha)^N 
      \left(\frac{1}{\alpha}\right)_k
      \left(\frac{2}{\alpha}\right)_k
      \left(\frac{\alpha{-}1}{\alpha}\right)_k
      \left(\frac{\alpha}{\alpha}\right)_k
  \end{align}
  which, when inserted into the appropriate hypergeometric function
  ${}_0F_\alpha$, yields
  \begin{align}
    \lleq{qpfa}
    {}_0F_\alpha\left[ 0,0,\ldots,0 \;;\;
    \frac{1}{\alpha},
    \frac{2}{\alpha},
    \ldots,
    \frac{\alpha{-}1}{\alpha},
    \frac{\alpha}{\alpha}
    \;;\; z \right]
    &= \sum_{k=0}^\infty \frac{z^k}{k!} \;\frac{(\alpha^\alpha)^k}{(k\alpha)!}
  \end{align}
  and with the general relation (see DLMF website)
  \begin{align}
    z\frac{d}{dz} {}_0F_p(\;;\;b_1,b_2,\ldots,b_q\;;\;z)
    &= \frac{z}{b_1b_2\cdots b_q}
      {}_0F_p(\;;\;b_1{+}1,b_2{+}1,\ldots,b_q{+}1\;;\;z)
  \end{align}
  we obtain, with the assignment
  \begin{align}
    z &= \lambda (\beta Q / \alpha)^\alpha \\
    \lleq{qpfe}
    p(Q\cond \lambda,\alpha,\beta)
      &= \frac{\alpha e^{-\lambda - \beta Q} }{Q} 
        z\frac{d}{dz} \;
        {}_0F_\alpha\left[ 0,0,\ldots,0 \;;\;
        \frac{1}{\alpha},
        \frac{2}{\alpha},
        \ldots,
        \frac{\alpha{-}1}{\alpha},
        \frac{\alpha}{\alpha}
        \;;\; z \right]
        \nonumber\\
      &= \frac{\alpha e^{-\lambda - \beta Q} }{Q} 
        \frac{z}{(\alpha-1)!\;\alpha^{\alpha-1}} \;
        {}_0F_\alpha\left[ 0,0,\ldots,0 \;;\;
        1{+}\frac{1}{\alpha},
        1{+}\frac{2}{\alpha},
        \ldots,
        1{+}\frac{\alpha{-}1}{\alpha},
        1{+}\frac{\alpha}{\alpha}
        \;;\; z \right]
        \nonumber\\
  \end{align}
  which on insertion of $z$ and simplification yields for integer $\alpha$
  \begin{equation}
    \lleq{qfp}
    \boxed{
      p(Q\cond \lambda,\alpha,\beta)
      = 
    \frac{\lambda\beta^\alpha Q^{\alpha-1} e^{-\lambda - \beta Q} }{(\alpha-1)!}\;
    {}_0F_\alpha\left[ \;;\;
        \frac{\alpha{+}1}{\alpha},
        \frac{\alpha{+}2}{\alpha},
        \ldots,
        \frac{2\alpha{-}1}{\alpha},
        2
        \;;\; \lambda\left(\frac{\beta Q}{\alpha}\right)^\alpha \;\right]
      }
  \end{equation}
  which includes the $I_1$ Bessel function for $\alpha=1$. Mathematica
  has the function ${}_pF_q$ available. Where it is not available, the
  above result should reduce, of course, to the series (\ref{gms}).
\end{itemize}

@@

\subsection{Moments of the luminosity prediction}\label{sec:predict-moments}

@@

\hans{H: I would agree to remove this section; I do not see value added.}
